Fail Faster

### 2025-03-25 paper/paper.tex
swarm plot labels
final edits

\documentclass[sigplan,screen,acmsmall,anonymous,review]{acmart}%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

@@

In Section~\ref{section:motiv}, we noted that generator combinators like \texttt{weighted\_union}
allocate lists in the hot path of the generator. Even though these lists are usually small---at most a few dozen elements---each allocation takes us closer to the next garbage collection\proposecut{, which is bad for performance}.

@@

In Figure~\ref{fig:codecps-and-final-gen}, following prior work \cite{kovacs24, carrette05}, we define
the type \texttt{'a CodeCps.t = 'z. ('a -> 'z code) -> 'z code}: a polymorphic continuation transformer with the result type
always in \texttt{code}.\footnote{This is an instance of the \emph{codensity} monad \cite{janis08}---a fact that deserves further investigation.}
The monad instance for this type is the standard instance for a CPS monad with polymorphic return type.
In prior work, this type is often referred to as the ``code generation'' monad.
This is because a value of type \texttt{('a code) CodeCps.t} is like an ``action'' that generates \texttt{code}:
\texttt{CodeCps.run} passes the continuation transformer the identity continuation to produce a \texttt{'a code}.
To avoid confusion with random data generators, we refer to this type as \texttt{CodeCps.t}.
Most importantly, the \texttt{CodeCps} type supports a function \texttt{let\_insert}, which, given \texttt{cx : 'a code},
let-binds \texttt{let x = .\~{}cx}, and then passes \texttt{.<x>.} to the
continuation.
\hg{I'm getting lost in the inline code again}

@@

In our setting, letting programmers define recursive generators as recursive functions is out of the question.
With staged programming, recursion must be handled with care: it is far too easy to accidentally recursively
define an infinite \texttt{code} value and have the program diverge at compile time\proposecut{, when trying to write
a \texttt{code} representing a recursive program.}
To this end, we develop a staged recursive generator combinator\footnote{
We actually provide a more general API that allows programmers to define \emph{parameterized}
recursive combinators, of type \texttt{'r code -> 'a code Gen.t}, for any type \texttt{'r}.
}, whose API is shown in Figure~\ref{fig:staged-recursive-generator}.
The recursion API consists of an opaque type \texttt{'a handle}, and a function \texttt{recurse} to perform recursive calls.
Programmers can then define recursive generators by \texttt{fixed\_point}, which ties the recursive knot.

>>>>

\documentclass[acmsmall,screen,review,anonymous]{acmart}
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    Bib\TeX}}}

@@

In Section~\ref{section:motiv}, we noted that generator combinators like \texttt{weighted\_union}
allocate lists in the hot path of the generator. Even though these lists are usually small---at most a few dozen elements---each allocation takes us closer to the next garbage collection, which is bad for performance.

@@

In Figure~\ref{fig:codecps-and-final-gen}, following prior work \cite{kovacs24, carrette05}, we define
the type \texttt{'a CodeCps.t = 'z. ('a -> 'z code) -> 'z code}: a polymorphic continuation transformer with the result type
always in \texttt{code}.\footnote{This is an instance of the \emph{codensity} monad \cite{janis08}---a fact that deserves further investigation.}
The monad instance for this type is the standard instance for a CPS monad with polymorphic return type.
In prior work, this type is often referred to as the ``code generation'' monad.
This is because a value of type \texttt{('a code) CodeCps.t} is like an ``action'' that generates \texttt{code}:
\texttt{CodeCps.run} passes the continuation transformer the identity continuation to produce a \texttt{'a code}.
To avoid confusion with random data generators, we refer to this type as \texttt{CodeCps.t}.
Most importantly, the \texttt{CodeCps} type supports a function \texttt{let\_insert}, which, given \texttt{cx : 'a code},
let-binds \texttt{let x = .\~{}cx}, and then passes \texttt{.<x>.} to the
continuation.
% \hg{I'm getting lost in the inline code again}

@@

In our setting, letting programmers define recursive generators as recursive functions is out of the question.
With staged programming, recursion must be handled with care: it is far too easy to accidentally recursively
define an infinite \texttt{code} value and have the program diverge at compile time, when trying to write
a \texttt{code} representing a recursive program.
To this end, we develop a staged recursive generator combinator\footnote{
We actually provide a more general API that allows programmers to define \emph{parameterized}
recursive combinators, of type \texttt{'r code -> 'a code Gen.t}, for any type \texttt{'r}.
}, whose API is shown in Figure~\ref{fig:staged-recursive-generator}.
The recursion API consists of an opaque type \texttt{'a handle}, and a function \texttt{recurse} to perform recursive calls.
Programmers can then define recursive generators by \texttt{fixed\_point}, which ties the recursive knot.
